---
title: 'NRSG 741 Homework 5'
author: "Alan Jones"
date: "March 15, 2017"
output:
  word_document: default
  pdf_document: default
  html_document:
    keep_md: yes
---

# Homework 5 - DUE March 15, 2017

For this homework, we'll work with the "Wong" dataset built in to the `car` package. The "Wong" data frame has 331 row and 7 columns. The observations are longitudinal data on recovery of IQ after comas of varying duration for 200 subjects. The data are from Wong, Monette, and Weiner (2001) and are for 200 patients who sustained traumatic brain injuries resulting in comas of varying duration. After awakening from their comas, patients were periodically administered a standard IQ test, but the average number of measurements per patient is small (331/200 = 1.7). *To get more info type `??Wong`.*

The 7 variables in the dataset are:

* `id`
    + patient ID number.
* `days`
    + number of days post coma at which IQs were measured.
* `duration`
    + duration of the coma in days.
* `sex`
    + a factor with levels Female and Male.
* `age`
    + in years at the time of injury.
* `piq`
    + performance (i.e., mathematical) IQ.
* `viq`
    + verbal IQ.

## Load dataset in from `car` package

```{r}
library(car)
data(Wong)

library(dplyr)
# add an age group variable
Wong$agegrp <- case_when(
  (Wong$age > 0 & Wong$age <= 10) ~ 1,
  (Wong$age > 10 & Wong$age <= 20) ~ 2,
  (Wong$age > 20 & Wong$age <= 30) ~ 3,
  (Wong$age > 30 & Wong$age <= 40) ~ 4,
  (Wong$age > 40 & Wong$age <= 50) ~ 5,
  (Wong$age > 50 & Wong$age <= 60) ~ 6,
  (Wong$age > 60 & Wong$age <= 70) ~ 7,
  (Wong$age > 70 & Wong$age <= 100) ~ 8)

# convert to factor, add code levels and labels
Wong$agegrp <- factor(Wong$agegrp,
                  levels = c(1,2,3,4,5,6,7,8),
                  labels = c("Ages 1-10",
                             "Ages 11-10",
                             "Ages 21-10",
                             "Ages 31-10",
                             "Ages 41-10",
                             "Ages 51-10",
                             "Ages 61-70",
                             "Ages 71-100"))
```

Using this dataset, and today's demos complete the following tasks:

1. Make a table of non-parametric statistics (median and IQR) for the number of days and duration grouped by `sex`. You'll be using `summarise()` from the `dplyr` package. For a given variable `x` you'll use `median(x, na.rm=TRUE)`, `quantile(x, 0.25, na.rm=TRUE)`, and `quantile(x, 0.75, na.rm=TRUE)`. Give the table a title using the `caption=` option and update the column names with something nice using the `col.names=` option in the `knitr::kable()` command. 
```{r}
Table1 <- Wong %>%
group_by(sex) %>%
summarise(firstquantiledays = quantile(days,0.25, na.rm=TRUE), mediandays = median(days, na.rm=TRUE), thirdrdquantiledays = quantile(days, 0.75, na.rm=TRUE), firstquantileduration =   quantile(duration, 0.25, na.rm=TRUE), medianduration = median(duration, na.rm=TRUE), thirdquantileduration = quantile(duration, 0.75, na.rm=TRUE))

knitr::kable(Table1, col.names=c("Sex", "1st Quantile Days", "Median Days", "3rd Quntile Days", "1st Quantile Duration", "Median Duration", "3rd Quantile Duration"), caption = "Summary Statistics of Days Post Coma and the Duration of Coma Grouped by Sex")

```


2.  Make a table of parametric statistics (mean and SD) for the performance outcomes `piq` and `viq` grouped by `sex`. Like the table above, you'll be using `summarise()` from the `dplyr` package. Now you'll use `mean(x, na.rm=TRUE)` and `sd(x, na.rm=TRUE)`. Give the table a title using the `caption=` option and update the column names with something nice using the `col.names=` option in the `knitr::kable()` command.
```{r}
Table2 <- Wong %>%
group_by(sex) %>%
summarise(avgperformance = mean(piq, na.rm=TRUE), sdperformance = sd(piq, na.rm=TRUE), avgverbal = mean(viq, na.rm=TRUE), sdverbal = sd(viq, na.rm=TRUE))

knitr::kable(Table2, col.names=c("Sex", "Mean Performance", "Performance SD", "Mean Verbal", "Verbal SD"), caption="Performance IQ and Verbal IQ Summary Statistics Grouped by Sex")

```


3. Make a table containing the frequencies and relative percentages for `agegrp`. Use the example we did in class to help guide you.
```{r}
#To add age group variable
Wong$agegrp <- case_when(
  (Wong$age > 0 & Wong$age <= 10) ~ 1,
  (Wong$age > 10 & Wong$age <= 20) ~ 2,
  (Wong$age > 20 & Wong$age <= 30) ~ 3,
  (Wong$age > 30 & Wong$age <= 40) ~ 4,
  (Wong$age > 40 & Wong$age <= 50) ~ 5,
  (Wong$age > 50 & Wong$age <= 60) ~ 6,
  (Wong$age > 60 & Wong$age <= 70) ~ 7,
  (Wong$age > 70 & Wong$age <= 100) ~ 8)

#To convert to factor and add code levels and labels
Wong$agegrp <- factor(Wong$agegrp, levels = c(1,2,3,4,5,6,7,8), labels = c("Ages 1-10", "Ages 11-10", "Ages 21-10", "Ages 31-10", "Ages 41-10", "Ages 51-10", "Ages 61-70", "Ages 71-100"))

#To find sample size or number of rows in dataset
ss <- length(Wong$agegrp)

#To group by each level of age group, find the raw counts and compute the relative %
Table3 <- Wong %>%
  group_by(agegrp) %>%
  summarise(freq = n(), pct = n()*100/ss)

#To make a summary table
knitr::kable(Table3, col.names = c("Age Group", "Frequency", "Percent"), caption="Frequency Table Sorted by Age Group")

```

4. Make a regression model (Model 1) for the performance IQ (`piq`) using `age` and `sex`. Put the regression model results into a table.
```{r}
m1 <- lm(piq~age + sex, data=Wong)
sm1 <- summary(m1)
knitr::kable(sm1$coefficients, caption="Model 1: Performance IQ vs Age and Sex")

```

5. Make a second regression model (Model 2) for performance IQ (`piq`) using `age` and `sex` plus `days` and `duration`. Put the regression model results into a table.
```{r}
m2 <- lm(piq~age+sex+days+duration, data=Wong)
sm2 <- summary(m2)
knitr::kable(sm2$coefficients, caption="Model 2: Performance IQ vs Age, Sex, Days and Duration")

```

6. Finally, make a table showing the results from the `anova()` command comparing Model 1 and Model 2 you made above using the example we did in class as a guide. 
```{r}
anova(m1,m2)
## Analysis of Variance Table
## 
## Model 1: piq ~ age + sex
## Model 2: piq ~ age + sex + days + duration
##   Res.Df   RSS Df Sum of Sq      F  Pr(>F)   
## 1    328 74968                               
## 2    326 72586  2    2381.9 5.3489 0.00518 **
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
## Analysis of Variance Table
## 
## Model 1: piq ~ age + sex
## Model 2: piq ~ age + sex + days + duration
##   Res.Df   RSS Df Sum of Sq      F  Pr(>F)   
## 1    328 74968                               
## 2    326 72586  2    2381.9 5.3489 0.00518 **
## ---
## Signif. codes:  0 '***' 0.001 '**' 0.01 '*' 0.05 '.' 0.1 ' ' 1
m1m2 <- anova(m1,m2)
row.names(m1m2) <- c("Model 1","Model 2")
knitr::kable(m1m2, caption = "Compare Model 1 and Model 2")

```

7. STUDENT CHOICE - pick either a `htmlwidget` from [http://gallery.htmlwidgets.org/](http://gallery.htmlwidgets.org/) or do a "flexdashboard" using the templates at [http://rmarkdown.rstudio.com/flexdashboard/](http://rmarkdown.rstudio.com/flexdashboard/) as a guide.

```{r setup, include=FALSE}
library(car)
data("Wong")
library(flexdashboard)
```
Column {data-width=650}
-----------------------------------------------------------------------

### Age in Wong dataset
```{r}
car::Wong
Wong$age
```
Column {data-width=350}
-----------------------------------------------------------------------

### Sex in Wong dataset
```{r}
car::Wong
Wong$sex
```
### Performance IQ in Wong dataset
```{r}
car::Wong
Wong$piq
```

### References

Wong, P. P., Monette, G., and Weiner, N. I. (2001) Mathematical models of cognitive recovery. Brain Injury, 15, 519–530.

Fox, J. (2016) Applied Regression Analysis and Generalized Linear Models, Third Edition. Sage.